Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-01-26 Manuscript/Manuscript.Rmd
Ampute impact data

# packages
library(tidyverse)
library(mice)
library(miceadds)
library(metamisc)
library(GJRM)
library(pan)

@@

```{r pop_dist, fig.cap = "Conditional distributions in the popularity data", echo=FALSE, message=FALSE, warning=FALSE}
plot_conditional(pop, x = "popteach", z = "popular", cluster = "class")
# TODO: check if up-side-down plot works
# TODO: check smoothing in geom_density function and make it the inverse of the sample size
# TODO: add functions to mice
# TODO: add title and informative legend
# TODO: add facets for some clusters
# or add propensity score distribution `is.na(popular) ~ .`
# TODO: make it average of cluster dens, not marginal
# alternatively add a quartile line around the density with geom_ribbon
```

@@

```{r impact, echo=FALSE}
# load data
data("impact")
# # descriptive statistics
# by(impact, impact$name, summary) 
# psych::describe(impact)[,c(2:5,8:9)]
# # missingness
# plot_md_pat(impact)
# TODO: rotate variable labels

@@

```{r hiv, echo=FALSE}
# data
data("hiv")
hiv <- hiv %>% 
  select(c("region","hiv", "age", "marital", "condom", "smoke")) %>% 
  mutate(region = as.integer(region))
# md pattern
plot_md_pat(hiv)
```

@@

The inclusion restriction variable should be a predictor of the missingness indicator for the variable of interest, but *not* of the actual value of the variable of interest. In this case, the data were simulated to adhere to this requirement. Namely, $\beta_{smoke}$ = `r mod_main$estimate`, 95% CI [`r paste0(mod_main$conf.low, ", ", mod_main$conf.high)`] for the analysis model (`formula = hiv ~ .`), and $\beta_{smoke}$ = `r mod_select$estimate`, 95% CI [`r paste0(mod_select$conf.low, ", ", mod_select$conf.high)`] for the selection model (`formula = is.na(hiv) ~ .`). This means the assumptions for the Heckman model are met.

>>>>

# packages
library(tidyverse)
library(mice)
library(miceadds)
# library(metamisc)
# library(GJRM)
# library(pan)

@@

```{r pop_dist, fig.cap = "Conditional distributions in the popularity data", echo=FALSE, message=FALSE, warning=FALSE}
plot_conditional(pop, x = "popteach", z = "popular", cluster = "class")
# TODO: check if up-side-down plot works
# TODO: check smoothing in geom_density function and make it the inverse of the sample size
# TODO: add functions to mice
# TODO: add title and informative legend
# TODO: add facets for some clusters
# or add propensity score distribution `is.na(popular) ~ .`
# TODO: make it average of cluster dens, not marginal
# alternatively add a quartile line around the density with geom_ribbon

## comments 24/1
# pop obs and pop miss in legend
# add to caption
# add fill
# overlap coeff??
# make densities proportional
# 
```

@@

```{r impact, echo=FALSE, fig.width=5}
# load data
data("impact", package = "metamisc")
impact_NA <- readRDS("../Data/impact_NA.RDS")
# # descriptive statistics
# by(impact, impact$name, summary) 
# psych::describe(impact)[,c(2:5,8:9)]
# # missingness
plot_md_pat(impact_NA)

# md pattern in first cluster
plot_md_pat(impact_NA[impact_NA$name == "TINT", ])
plot_md_pat(impact_NA[impact_NA$name == "APOE", ])
# TODO: rotate variable labels

@@

```{r hiv, echo=FALSE}
# data
data("hiv", package = "GJRM")
hiv <- hiv %>% 
  select(c("region","hiv", "age", "marital", "condom", "smoke")) %>% 
  mutate(region = as.integer(region))
# md pattern
plot_md_pat(hiv)
```

@@

The inclusion restriction variable should be a predictor of the the actual value of the variable of interest, but *not* of missingness indicator for the variable of interest. In this case, the data were simulated to adhere to this requirement. Namely, $\beta_{smoke}$ = `r mod_main$estimate`, 95% CI [`r paste0(mod_main$conf.low, ", ", mod_main$conf.high)`] for the analysis model (`formula = hiv ~ .`), and $\beta_{smoke}$ = `r mod_select$estimate`, 95% CI [`r paste0(mod_select$conf.low, ", ", mod_select$conf.high)`] for the selection model (`formula = is.na(hiv) ~ .`). This means the assumptions for the Heckman model are met.

### 2022-01-26 Manuscript/Manuscript.tex
Ampute impact data

@@

The inclusion restriction variable should be a predictor of the
missingness indicator for the variable of interest, but \emph{not} of
the actual value of the variable of interest. In this case, the data
were simulated to adhere to this requirement. Namely, \(\beta_{smoke}\)
= -0.064, 95\% CI {[}-0.256, 0.126{]} for the analysis model
(\texttt{formula\ =\ hiv\ \textasciitilde{}\ .}), and \(\beta_{smoke}\)
= -0.265, 95\% CI {[}-0.422, -0.11{]} for the selection model
(\texttt{formula\ =\ is.na(hiv)\ \textasciitilde{}\ .}). This means the
assumptions for the Heckman model are met.

>>>>

\begin{center}\includegraphics{Manuscript_files/figure-latex/impact-1} \end{center}

\begin{center}\includegraphics{Manuscript_files/figure-latex/impact-2} \end{center}

\begin{center}\includegraphics{Manuscript_files/figure-latex/impact-3} \end{center}

@@

The inclusion restriction variable should be a predictor of the the
actual value of the variable of interest, but \emph{not} of missingness
indicator for the variable of interest. In this case, the data were
simulated to adhere to this requirement. Namely, \(\beta_{smoke}\) =
-0.064, 95\% CI {[}-0.256, 0.126{]} for the analysis model
(\texttt{formula\ =\ hiv\ \textasciitilde{}\ .}), and \(\beta_{smoke}\)
= -0.265, 95\% CI {[}-0.422, -0.11{]} for the selection model
(\texttt{formula\ =\ is.na(hiv)\ \textasciitilde{}\ .}). This means the
assumptions for the Heckman model are met.
